---
title: "Analysis"
output:  
  html_document:
    number_sections: true
    toc: true  
    collapsed: false
    toc_float: true
    smooth_scroll: false
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
# packages
library(ez)
library(reshape2)
library(reshape)
library(ggplot2)
library(plyr)
library(pastecs)
library(ez)
library(data.table)
library(tidyverse) 

library(showtext)
library(readr)


font_add_google("Poppins", "Poppins")
font_add_google("Roboto Mono", "Roboto Mono")
showtext_auto()
```


# Data loadind and cleaning 

Load the data 

```{r}
tdata <- read_csv("exp_data.csv")
```


Exclude data from subjects who failed the check questions (attention, learning, comprehension, etc.)

```{r}
# Exclude subjects who failed the control questions 
tdata <- subset(tdata, confirmations_correct == "correct")
tdata <- subset(tdata, intro_check_correct == "correct")
tdata <- subset(tdata, learn_check_correct == "correct")
tdata <- subset(tdata, learn_check_structure_correct == "correct")
tdata <- subset(tdata, timing_check_correct == "correct")
```


```{r}
table(tdata$structure, tdata$first_object)
```
Table shows that the admired sample sizes (N = 150, n = 25 per group) were reached.


# Subject demographics

```{r}
# demographics 

min(tdata$age)
max(tdata$age)
mean(tdata$age)
sd(tdata$age)

# 1 = male, 2 = female, 3 = other
table(tdata$gender)
```

1 = male, 2 = female, 3 = non-binary, 4 = prefer not to say


# Data preparations 


Factorize:

```{r}
tdata$structure <- factor(tdata$structure, levels = c("stable", "unstable", "reversible"), 
                          labels = c("one-shot", "demarcating and irreversible", "demarcating and reversible"))
```

Relabel rating columns for the two causes according to "first object" (so that they are called first_obj_rating and second_obj_rating):

```{r}
left_first <- subset(tdata, first_object == "left")
colnames(left_first)[16] ="first_obj_caused_rating"
colnames(left_first)[17] ="second_obj_caused_rating"
colnames(left_first)[18] ="first_obj_maintained_rating"
colnames(left_first)[19] ="second_obj_maintained_rating"

right_first <- subset(tdata, first_object == "right")
colnames(right_first)[17] ="first_obj_caused_rating"
colnames(right_first)[16] ="second_obj_caused_rating"
colnames(right_first)[19] ="first_obj_maintained_rating"
colnames(right_first)[18] ="second_obj_maintained_rating"

right_first <- right_first[c("subj_code", "condition", "desktop_conf", "attent_conf", "confirmations_correct", "intro_check", "intro_check_correct",
                             "structure", "first_object", "learn_check_left", "learn_check_left_correct", "learn_check_right", "learn_check_right_correct",
                             "learn_check_structure", "learn_check_structure_correct",
                             "first_obj_caused_rating", "second_obj_caused_rating", "first_obj_maintained_rating", "second_obj_maintained_rating",
                             "timing_check", "timing_check_correct", "recorded_at", 
                             "age", "gender", "tech_issues", 
                             "learn_check_correct")]


names(left_first) %in% names(right_first)

tdata <- rbind(left_first, right_first)
```

Make a subset containing only the columns relevant for analyses and turn into long format:

```{r}
# Subset: 
tdata_sub <- subset(tdata, select = c("subj_code","structure","first_obj_caused_rating", "second_obj_caused_rating",
                                      "first_obj_maintained_rating", "second_obj_maintained_rating"))

# into long format:
library(tidyr)
tdata_long <- gather(tdata_sub, entity, rating, 3:6)


# factorize entity 
tdata_long$entity <- factor(tdata_long$entity, levels = c("first_obj_caused_rating", "second_obj_caused_rating", 
                                                          "first_obj_maintained_rating", "second_obj_maintained_rating"), 
                            labels = c("first \ntriggered", "second \ntriggered","first \nsustained", "second \nsustained"))

tdata_long$structure <- factor(tdata_long$structure, levels = c("one-shot", "demarcating and irreversible", "demarcating and reversible"))
```



# Graphs and summaries


```{r}
myTheme <- theme(plot.title = element_text(face="bold", size = 22),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        axis.text.x = element_text(size = 14, angle = 0), 
        axis.text.y = element_text(size = 16, angle = 0),
        legend.text = element_text(size = 18),
        legend.title = element_text(face = "bold", size = 18),
        strip.text.x = element_text(size = 18),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        panel.background = element_blank(), 
        axis.line.x = element_line(colour = "black"), 
        axis.line.y = element_line(colour = "black"),
        axis.text = element_text(colour ="black"), 
        axis.ticks = element_line(colour ="black"))


tdata_sub <- tdata_long


library(see)
## first, turn sID into a factor
tdata_sub$subj_code <- factor(tdata_sub$subj_code)

pd <- position_dodge(width = 0.3)

tdata_sub$valueJitter <- jitter(tdata_sub$rating, factor = 0.01, amount = 0.004)

theme_set(theme_light(base_size = 20, base_family = "Poppins"))


g <- ggplot(tdata_sub, aes(x = entity, y = valueJitter)) +
  guides(fill=FALSE)+
  facet_grid( ~ structure)+
  scale_y_continuous(limits = c(-0.3, 10.3), breaks=seq(0, 10, 1), expand = c(0,0)) +
  geom_jitter(aes(color = structure), alpha = 0.5, width = 0.15, height = 0.2) +
  stat_summary(aes(y = rating, group=1), fun.data = mean_cl_boot, 
               geom = "errorbar", width = 0, size = 1) +
  stat_summary(aes(y = rating, group=1, color = structure), fun.y=mean, geom="line", 
               shape = 22, size = 1.5, alpha = .7)+
  stat_summary(aes(y = rating, group=1, fill = structure), fun.y=mean, geom="point", 
               color = "black", shape = 22, size = 3, group=1, alpha = 1)+
  stat_summary(aes(y = rating,group=1), fun.y=median, geom="point", color = "black", shape = 3, size = 4, 
               group=1, alpha = 1, position = position_dodge(width = 0.5))+
  labs(x = "Causal Statement", y = "Agreement Rating") +
  scale_color_manual(name = "Strength",values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  scale_fill_manual(name = "Strength",values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  theme(legend.position = "none")+
  myTheme+
  theme(panel.grid.major = element_line(color = "lightgrey",
                                        size = 0.5,
                                        linetype = 'dotted'))+
  stat_summary(aes(label=round(after_stat(y),2)), fun.y=mean, geom="text", size=4,
             vjust = 0, hjust = -0.5)+
  annotate("text", x = 0.5, y = 2.2, label = c("completely inadequate"), angle = 90)+
  annotate("text", x = 0.5, y = 8.0, label = c("completely adequate"), angle = 90)

g

ggsave("results_means_mainDV.svg",width=17,height=5)
ggsave("results_means_mainDV.pdf",width=17,height=5)

```

Extract from the plot a list of the means and their 95% CIs: 

```{r}
# Build the ggplot object
gb <- ggplot_build(g)

# Extract the data from the stat_summary layer
ci_data <- gb$data[[2]] # stat_summary is the 2nd layer

# Extract the lower and upper CI boundaries
lower_ci <- ci_data$ymin
upper_ci <- ci_data$ymax

# Extract the means
means <- ci_data$y

# Combine into a data frame (optional)
ci_df <- data.frame(
  group = ci_data$x,
  mean = means,
  lower_ci = lower_ci,
  upper_ci = upper_ci
)

# Print the data frame
print(ci_df)
```



```{r}
library(ggridges)
g2 <- ggplot(tdata_long, aes(x = rating, y = entity, fill = structure)) +
  facet_grid( ~ structure)+
  scale_x_continuous(breaks = seq(0, 10, 1))+
  geom_density_ridges(alpha = 0.5)+
   #stat_summary(aes(x = rating_rec), fun.x=mean, geom="point", 
  #             color = "black", shape = 22, size = 2, group=1, alpha = 1)+
  scale_fill_manual(values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  #scale_fill_viridis_c(name = "Explanation \nRating", option = "C", breaks=c(-5,0,5), labels=c("narrow scope", "no preference", "broad scope"))+
  labs(x = "Agreement Rating", y = "Causal Statement") +
  scale_y_discrete(limits=rev)+
  myTheme+
  theme_light(base_family = "Poppins", base_size = 20)+
  theme(panel.grid = element_blank(), axis.text = element_text(colour ="black"))+
  theme(legend.position="none",
        legend.title=element_blank(),legend.key.width = unit(1.95, 'cm'))+
  theme(axis.text.y = element_text(size = 14, angle = 0))+
  annotate("text", y = 0.7, x = 0, label = c("completely inadequate"), angle = 0)+
  annotate("text", y = 0.7, x = 10, label = c("completely adequate"), angle = 0)

g2

ggsave("results_dist.svg",width=17,height=5)
ggsave("results_dist.pdf",width=17,height=5)
```


```{r}
library("ggpubr")

ggarrange(
  g, g2, labels = c("A", "B"),
  common.legend = TRUE, legend = "none", 
  font.label = list(size = 16, color = "black", family = "Arial", face = "bold"), 
  nrow = 2,
  ncol = 1
  ) -> c_plot

c_plot

ggsave("res_squids_02.svg",width=16,height=12)
ggsave("res_squids_02.pdf",width=16,height=12)
```


# Analyses

## Mixed ANOVA 

```{r}
library(afex)
library(emmeans)

a1 <- aov_car(rating ~ structure*entity + Error(subj_code/entity), tdata_long, anova_table = list(es = "pes"))
a1

```
Predicted interaction effect was significant-


Visualize also the obtained main effect of "structure" (reversibility)

```{r}
tdata_sub <- tdata_long


library(see)
## first, turn sID into a factor
tdata_sub$subj_code <- factor(tdata_sub$subj_code)

pd <- position_dodge(width = 0.3)

tdata_sub$valueJitter <- jitter(tdata_sub$rating, factor = 0.01, amount = 0.004)

theme_set(theme_light(base_size = 20, base_family = "Poppins"))


g <- ggplot(tdata_sub, aes(x = structure, y = valueJitter)) +
  guides(fill=FALSE)+

  scale_y_continuous(limits = c(-0.3, 10.3), breaks=seq(0, 10, 1), expand = c(0,0)) +
  geom_jitter(aes(color = structure), alpha = 0.5, width = 0.15, height = 0.2) +
  stat_summary(aes(y = rating, group=1), fun.data = mean_cl_boot, 
               geom = "errorbar", width = 0, size = 1) +
  stat_summary(aes(y = rating, group=1, fill = structure), fun.y=mean, geom="point", 
               color = "black", shape = 22, size = 3, group=1, alpha = 1)+
  stat_summary(aes(y = rating,group=1), fun.y=median, geom="point", color = "black", shape = 3, size = 4, 
               group=1, alpha = 1, position = position_dodge(width = 0.5))+
  labs(x = "Nature of the effect", y = "Agreement Rating") +
  scale_color_manual(name = "Strength",values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  scale_fill_manual(name = "Strength",values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  theme(legend.position = "none")+
  myTheme+
  theme(panel.grid.major = element_line(color = "lightgrey",
                                        size = 0.5,
                                        linetype = 'dotted'))+
  stat_summary(aes(label=round(after_stat(y),2)), fun.y=mean, geom="text", size=5,
             vjust = -6)+
  annotate("text", x = 0.5, y = 2, label = c("completely disagree"), angle = 90)+
  annotate("text", x = 0.5, y = 8.2, label = c("completely agree"), angle = 90)

g

```

Plot main effect of entity (causal statement):

```{r}

tdata_sub <- tdata_long


library(see)
## first, turn sID into a factor
tdata_sub$subj_code <- factor(tdata_sub$subj_code)

pd <- position_dodge(width = 0.3)

tdata_sub$valueJitter <- jitter(tdata_sub$rating, factor = 0.01, amount = 0.004)

theme_set(theme_light(base_size = 20, base_family = "Poppins"))

g <- ggplot(tdata_sub, aes(x = entity, y = valueJitter)) +
  guides(fill=FALSE)+
  scale_y_continuous(limits = c(-0.3, 10.3), breaks=seq(0, 10, 1), expand = c(0,0)) +
  geom_jitter(aes(color = entity), alpha = 0.5, width = 0.15, height = 0.2) +
  stat_summary(aes(y = rating, group=1), fun.data = mean_cl_boot, 
               geom = "errorbar", width = 0, size = 1) +
  stat_summary(aes(y = rating, group=1, fill = entity), fun.y=mean, geom="point", 
               color = "black", shape = 22, size = 3, group=1, alpha = 1)+
  stat_summary(aes(y = rating,group=1), fun.y=median, geom="point", color = "black", shape = 3, size = 4, 
               group=1, alpha = 1, position = position_dodge(width = 0.5))+
  labs(x = "Causal Statement", y = "Agreement Rating") +
  scale_color_manual(name = "Strength",values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  scale_fill_manual(name = "Strength",values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  theme(legend.position = "none")+
  annotate("text", x = 0.5, y = 1.7, label = c("completely disagree"), angle = 90)+
  annotate("text", x = 0.5, y = 6.7, label = c("completely agree"), angle = 90)+
  myTheme+
  theme(panel.grid.major = element_line(color = "lightgrey",
                                        size = 0.5,
                                        linetype = 'dotted'))+
  stat_summary(aes(label=round(after_stat(y),2)), fun.y=mean, geom="text", size=5,
             vjust = -6)

g


```


## Contrasts 



```{r}

library(lsmeans)
# means

ls2 <- lsmeans(a1, c("structure", "entity")) 
ls2

contrasts <- emmeans(a1, ~ entity|structure)
s <- pairs(contrasts, adjust = "none") 


s
confint(s, level = 0.95)
```

The planned contrasts from the list above are " first..caused - second..caused" in both "irreversible" and "reversible". Both are significant.

Get effect standardized effect sizes for the contrasts

First get the SDs that need to be pooled (and also the correlations between measures for the within-comparisons)

```{r}
library(dplyr)

tdata_long %>%
  group_by(structure, entity) %>%
  summarise_at(vars(rating), list(name=sd))
```



The planned contrasts compare the maintainer ratings for the second cause in "stable", "unstable", and "reversible". It must be looked up from the long list of contrasts below: 

```{r}
contrasts <- emmeans(a1, ~ entity*structure)
s <- pairs(contrasts, adjust = "none") 


s
confint(s, level = 0.95)
```

The relevant contrasts are: "second..maintained stable - second..maintained unstable", which is significant. 

The relevant contrasts are: "second..maintained stable - second..maintained reversible", which is significant. 


Get effect size: 




```{r}
# using the functions from the MOTE package (see https://matthewbjane.quarto.pub/guide-to-effect-sizes-and-confidence-intervals/Standardized-Mean-Differences.html#sec-repeated-measures-drm)

library(MOTE)

#Means: 
# stable effect     second..maintained   3.10 0.389 147    2.332     3.87
# unstable effect   second..maintained   9.52 0.389 147    8.752    10.29
# reversible effect second..maintained   9.00 0.389 147    8.232     9.77
 
#SDs: 
# stable effect	second \nmaintained	3.9136079		
# unstable effect	second \nmaintained	1.3284946		
# reversible effect	second \nmaintained	2.3647065	

# First stable vs. unstable

stats <- d.ind.t(
  m1 = 9.52,
  m2 = 3.10,
  sd1 = 1.3284946,
  sd2 = 3.9136079,
  n1 = 50,
  n2 = 50,
  a = 0.05
)

stats$estimate
```

```{r}
# using the functions from the MOTE package (see https://matthewbjane.quarto.pub/guide-to-effect-sizes-and-confidence-intervals/Standardized-Mean-Differences.html#sec-repeated-measures-drm)

library(MOTE)

#Means: 
# stable effect     second..maintained   3.10 0.389 147    2.332     3.87
# unstable effect   second..maintained   9.52 0.389 147    8.752    10.29
# reversible effect second..maintained   9.00 0.389 147    8.232     9.77
 
#SDs: 
# stable effect	second \nmaintained	3.9136079		
# unstable effect	second \nmaintained	1.3284946		
# reversible effect	second \nmaintained	2.3647065	

# First stable vs. reversible

stats <- d.ind.t(
  m1 = 9.00,
  m2 = 3.10,
  sd1 = 2.3647065	,
  sd2 = 3.9136079,
  n1 = 50,
  n2 = 50,
  a = 0.05
)

stats$estimate
```


# Test only the unstable and reversible effect conditions 

```{r}
library(afex)
library(emmeans)

tdata_rev_unst <- subset(tdata_long, structure != "stable effect")

a2 <- aov_car(rating ~ structure*entity + Error(subj_code/entity), tdata_rev_unst, anova_table = list(es = "pes"))
a2

```
