---
title: "Analysis"
output:  
  html_document:
    number_sections: true
    toc: true  
    collapsed: false
    toc_float: true
    smooth_scroll: false
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
# packages
library(ez)
library(reshape2)
library(reshape)
library(ggplot2)
library(plyr)
library(pastecs)
library(ez)
library(data.table)
library(tidyverse) 

library(showtext)
library(readr)


font_add_google("Poppins", "Poppins")
font_add_google("Roboto Mono", "Roboto Mono")
showtext_auto()
```


```{r}
tdata <- read_csv("exp_data.csv")
```


```{r}
# Exclude subjects who failed the control questions 
tdata <- subset(tdata, confirmations_correct == "correct")
tdata <- subset(tdata, intro_check_correct == "correct")
tdata <- subset(tdata, learning_check_correct == "correct")
tdata <- subset(tdata, learn_check_structure_correct == "correct")
tdata <- subset(tdata, timing_check_correct == "correct")
```


```{r}
table(tdata$structure, tdata$first_object)
```

# Subject demographics

```{r}
# demographics 

min(tdata$age)
max(tdata$age)
mean(tdata$age)
sd(tdata$age)

# 1 = male, 2 = female, 3 = other
table(tdata$gender)
```

1 = male, 2 = female, 3 = non-binary, 4 = prefer not to say


# Data preparations 


Factorize:

```{r}
tdata$structure <- factor(tdata$structure, levels = c("stable", "unstable"), labels = c("one-shot", "demarcation"))
```

Relabel rating columns for red and yellow entity according to "first object" (so that they are called first_obj_rating and second_obj_rating):

```{r}
left_first <- subset(tdata, first_object == "left")
colnames(left_first)[11] ="first_obj_caused_rating"
colnames(left_first)[12] ="second_obj_caused_rating"
colnames(left_first)[13] ="first_obj_sustained_rating"
colnames(left_first)[14] ="second_obj_sustained_rating"

right_first <- subset(tdata, first_object == "right")
colnames(right_first)[12] ="first_obj_caused_rating"
colnames(right_first)[11] ="second_obj_caused_rating"
colnames(right_first)[14] ="first_obj_sustained_rating"
colnames(right_first)[13] ="second_obj_sustained_rating"

right_first <- right_first[c("subj_code", "condition", "desktop_conf", "attent_conf", "intro_check", 
                             "structure", "first_object", "learn_check_left", "learn_check_right", "learn_check_structure",
                             "first_obj_caused_rating", "second_obj_caused_rating", "first_obj_sustained_rating", "second_obj_sustained_rating",
                             "timing_check", "recorded_at", 
                             "age", "gender", "tech_issues", 
                             "confirmations_correct", "intro_check_correct",
                             "learning_check_left_correct", "learning_check_right_correct", "learning_check_correct",
                             "learn_check_structure_correct", "timing_check_correct")]


names(left_first) %in% names(right_first)

tdata <- rbind(left_first, right_first)
```

Make a subset containing only the columns relevant for analyses and turn into long format:

```{r}
# Subset: 
tdata_sub <- subset(tdata, select = c("subj_code","structure","first_obj_caused_rating", "second_obj_caused_rating",
                                      "first_obj_sustained_rating", "second_obj_sustained_rating"))

# into long format:
library(tidyr)
tdata_long <- gather(tdata_sub, entity, rating, 3:6)


# factorize entity 
tdata_long$entity <- factor(tdata_long$entity, levels = c("first_obj_caused_rating", "second_obj_caused_rating", 
                                                          "first_obj_sustained_rating", "second_obj_sustained_rating"), 
                            labels = c("first \ntriggered", "second \ntriggered","first \nsustained", "second \nsustained"))
```



# Graphs and summaries


```{r}
myTheme <- theme(plot.title = element_text(face="bold", size = 22),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        axis.text.x = element_text(size = 14, angle = 0), 
        axis.text.y = element_text(size = 16, angle = 0),
        legend.text = element_text(size = 18),
        legend.title = element_text(face = "bold", size = 18),
        strip.text.x = element_text(size = 18),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        panel.background = element_blank(), 
        axis.line.x = element_line(colour = "black"), 
        axis.line.y = element_line(colour = "black"),
        axis.text = element_text(colour ="black"), 
        axis.ticks = element_line(colour ="black"))


tdata_sub <- tdata_long


library(see)
## first, turn sID into a factor
tdata_sub$subj_code <- factor(tdata_sub$subj_code)

pd <- position_dodge(width = 0.3)

tdata_sub$valueJitter <- jitter(tdata_sub$rating, factor = 0.01, amount = 0.004)

theme_set(theme_light(base_size = 20, base_family = "Poppins"))


g <- ggplot(tdata_sub, aes(x = entity, y = valueJitter)) +
  guides(fill=FALSE)+
  facet_grid( ~ structure)+
  scale_y_continuous(limits = c(-0.3, 10.3), breaks=seq(0, 10, 1), expand = c(0,0)) +
  geom_jitter(aes(color = structure), alpha = 0.5, width = 0.15, height = 0.2) +
  stat_summary(aes(y = rating, group=1), fun.data = mean_cl_boot, 
               geom = "errorbar", width = 0, size = 1) +
  stat_summary(aes(y = rating, group=1, color = structure), fun.y=mean, geom="line", 
               shape = 22, size = 1.5, alpha = .7)+
  stat_summary(aes(y = rating, group=1, fill = structure), fun.y=mean, geom="point", 
               color = "black", shape = 22, size = 3, group=1, alpha = 1)+
  stat_summary(aes(y = rating,group=1), fun.y=median, geom="point", color = "black", shape = 3, size = 4, 
               group=1, alpha = 1, position = position_dodge(width = 0.5))+
  labs(x = "Causal Statement", y = "Agreement Rating") +
  scale_color_manual(name = "Strength",values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  scale_fill_manual(name = "Strength",values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  theme(legend.position = "none")+
  myTheme+
  theme(panel.grid.major = element_line(color = "lightgrey",
                                        size = 0.5,
                                        linetype = 'dotted'))+
  stat_summary(aes(label=round(after_stat(y),2)), fun.y=mean, geom="text", size=4,
             vjust = 0, hjust = -0.5)+
  annotate("text", x = 0.5, y = 2.2, label = c("completely inadequate"), angle = 90)+
  annotate("text", x = 0.5, y = 8.0, label = c("completely adequate"), angle = 90)

g

ggsave("results_means_mainDV.svg",width=12,height=5)
ggsave("results_means_mainDV.pdf",width=12,height=5)

```
Extract from the plot a list of the means and their 95% CIs: 

```{r}
# Build the ggplot object
gb <- ggplot_build(g)

# Extract the data from the stat_summary layer
ci_data <- gb$data[[2]] # stat_summary is the 2nd layer

# Extract the lower and upper CI boundaries
lower_ci <- ci_data$ymin
upper_ci <- ci_data$ymax

# Extract the means
means <- ci_data$y

# Combine into a data frame (optional)
ci_df <- data.frame(
  group = ci_data$x,
  mean = means,
  lower_ci = lower_ci,
  upper_ci = upper_ci
)

# Print the data frame
print(ci_df)
```



```{r}
library(ggridges)
g2 <- ggplot(tdata_long, aes(x = rating, y = entity, fill = structure)) +
  facet_grid( ~ structure)+
  scale_x_continuous(breaks = seq(0, 10, 1))+
  geom_density_ridges(alpha = 0.5)+
   #stat_summary(aes(x = rating_rec), fun.x=mean, geom="point", 
  #             color = "black", shape = 22, size = 2, group=1, alpha = 1)+
  scale_fill_manual(values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  #scale_fill_viridis_c(name = "Explanation \nRating", option = "C", breaks=c(-5,0,5), labels=c("narrow scope", "no preference", "broad scope"))+
  labs(x = "Agreement Rating", y = "Causal Statement") +
  scale_y_discrete(limits=rev)+
  myTheme+
  theme_light(base_family = "Poppins", base_size = 20)+
  theme(panel.grid = element_blank(), axis.text = element_text(colour ="black"))+
  theme(legend.position="none",
        legend.title=element_blank(),legend.key.width = unit(1.95, 'cm'))+
  theme(axis.text.y = element_text(size = 14, angle = 0))+
  annotate("text", y = 0.7, x = 0, label = c("completely inadequate"), angle = 0)+
  annotate("text", y = 0.7, x = 10, label = c("completely adequate"), angle = 0)

g2

ggsave("results_dist.svg",width=12,height=5)
ggsave("results_dist.pdf",width=12,height=5)
```
```{r}
library("ggpubr")

ggarrange(
  g, g2, labels = c("A", "B"),
  common.legend = TRUE, legend = "none", 
  font.label = list(size = 16, color = "black", family = "Arial", face = "bold"), 
  nrow = 2,
  ncol = 1
  ) -> c_plot

c_plot

ggsave("res_squids_01.svg",width=11,height=12)
ggsave("res_squids_01.pdf",width=11,height=12)
```



# Analyses

## Mixed ANOVA 

```{r}
library(afex)
library(emmeans)

a1 <- aov_car(rating ~ structure*entity + Error(subj_code/entity), tdata_long, anova_table = list(es = "pes"))
a1

```
Predicted interaction effect was significant-


Visualize also the obtained main effect of "structure" (reversibility)

```{r}
tdata_sub <- tdata_long


library(see)
## first, turn sID into a factor
tdata_sub$subj_code <- factor(tdata_sub$subj_code)

pd <- position_dodge(width = 0.3)

tdata_sub$valueJitter <- jitter(tdata_sub$rating, factor = 0.01, amount = 0.004)

theme_set(theme_light(base_size = 20, base_family = "Poppins"))


g <- ggplot(tdata_sub, aes(x = structure, y = valueJitter)) +
  guides(fill=FALSE)+

  scale_y_continuous(limits = c(-0.3, 10.3), breaks=seq(0, 10, 1), expand = c(0,0)) +
  geom_jitter(aes(color = structure), alpha = 0.5, width = 0.15, height = 0.2) +
  stat_summary(aes(y = rating, group=1), fun.data = mean_cl_boot, 
               geom = "errorbar", width = 0, size = 1) +
  stat_summary(aes(y = rating, group=1, fill = structure), fun.y=mean, geom="point", 
               color = "black", shape = 22, size = 3, group=1, alpha = 1)+
  stat_summary(aes(y = rating,group=1), fun.y=median, geom="point", color = "black", shape = 3, size = 4, 
               group=1, alpha = 1, position = position_dodge(width = 0.5))+
  labs(x = "Nature of the effect", y = "Agreement Rating") +
  scale_color_manual(name = "Strength",values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  scale_fill_manual(name = "Strength",values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  theme(legend.position = "none")+
  myTheme+
  theme(panel.grid.major = element_line(color = "lightgrey",
                                        size = 0.5,
                                        linetype = 'dotted'))+
  stat_summary(aes(label=round(after_stat(y),2)), fun.y=mean, geom="text", size=5,
             vjust = -6)+
  annotate("text", x = 0.5, y = 2, label = c("completely disagree"), angle = 90)+
  annotate("text", x = 0.5, y = 8.2, label = c("completely agree"), angle = 90)

g

```

Plot main effect of entity (causal statement):

```{r}

tdata_sub <- tdata_long


library(see)
## first, turn sID into a factor
tdata_sub$subj_code <- factor(tdata_sub$subj_code)

pd <- position_dodge(width = 0.3)

tdata_sub$valueJitter <- jitter(tdata_sub$rating, factor = 0.01, amount = 0.004)

theme_set(theme_light(base_size = 20, base_family = "Poppins"))

g <- ggplot(tdata_sub, aes(x = entity, y = valueJitter)) +
  guides(fill=FALSE)+
  scale_y_continuous(limits = c(-0.3, 10.3), breaks=seq(0, 10, 1), expand = c(0,0)) +
  geom_jitter(aes(color = entity), alpha = 0.5, width = 0.15, height = 0.2) +
  stat_summary(aes(y = rating, group=1), fun.data = mean_cl_boot, 
               geom = "errorbar", width = 0, size = 1) +
  stat_summary(aes(y = rating, group=1, fill = entity), fun.y=mean, geom="point", 
               color = "black", shape = 22, size = 3, group=1, alpha = 1)+
  stat_summary(aes(y = rating,group=1), fun.y=median, geom="point", color = "black", shape = 3, size = 4, 
               group=1, alpha = 1, position = position_dodge(width = 0.5))+
  labs(x = "Causal Statement", y = "Agreement Rating") +
  scale_color_manual(name = "Strength",values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  scale_fill_manual(name = "Strength",values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  theme(legend.position = "none")+
  annotate("text", x = 0.5, y = 1.7, label = c("completely disagree"), angle = 90)+
  annotate("text", x = 0.5, y = 6.7, label = c("completely agree"), angle = 90)+
  myTheme+
  theme(panel.grid.major = element_line(color = "lightgrey",
                                        size = 0.5,
                                        linetype = 'dotted'))+
  stat_summary(aes(label=round(after_stat(y),2)), fun.y=mean, geom="text", size=5,
             vjust = -6)

g


```


## Contrasts 



```{r}

library(lsmeans)
# means

ls2 <- lsmeans(a1, c("structure", "entity")) 
ls2

contrasts <- emmeans(a1, ~ entity|structure)
s <- pairs(contrasts, adjust = "none") 


s
confint(s, level = 0.95)
```

The planned contrasts from the list above are " first..caused - second..caused" in both "irreversible" and "reversible". Both are significant.

Get effect standardized effect sizes for the contrasts

First get the SDs that need to be pooled (and also the correlations between measures for the within-comparisons)

```{r}
library(dplyr)

tdata_long %>%
  group_by(structure, entity) %>%
  summarise_at(vars(rating), list(name=sd))
```


```{r}
library(dplyr)

tdata_stable <- subset(tdata, structure == "one-shot")
tdata_unstable <- subset(tdata, structure == "demarcation")

tdata_stable %>%
  group_by(structure) %>%
  summarize(cor=cor(first_obj_caused_rating, second_obj_caused_rating))

tdata_unstable %>%
  group_by(structure) %>%
  summarize(cor=cor(first_obj_caused_rating, second_obj_caused_rating))

```

```{r}
# using the functions from the MOTE package (see https://matthewbjane.quarto.pub/guide-to-effect-sizes-and-confidence-intervals/Standardized-Mean-Differences.html#sec-repeated-measures-drm)

library(MOTE)


# stable condition

# structure       entity             lsmean    SE df lower.CL upper.CL
# stable effect   first..caused        9.84 0.224 98   9.3946    10.29
# stable effect   second..caused       0.86 0.391 98   0.0833     1.64

# sds
#stable effect	first \ncaused	0.7384485		
#stable effect	second \ncaused	1.9795897	

#Cor
#irreversible	-0.1470522		
	

stats <- d.dep.t.rm(
  m1 = 9.84,
  m2 = 0.86,
  sd1 = 0.7384485	,
  sd2 = 1.9795897	,
  n = 50,
  a = 0.05,
  r = -0.3367331	
)

stats$estimate
```

```{r}
# unstable condition 

#structure    entity             lsmean    SE df lower.CL upper.CL
# unstable effect first..caused        9.28 0.224 98   8.8346     9.73
# unstable effect second..caused       2.22 0.391 98   1.4433     3.00

# sds 
# unstable effect	first \ncaused	2.1192991		
# unstable effect	second \ncaused	3.3763282	

#Cor
#unstable	-0.06582681	


stats <- d.dep.t.rm(
  m1 = 9.28,
  m2 = 2.22,
  sd1 = 2.1192991		,
  sd2 = 3.3763282,
  n = 50,
  a = 0.05,
  r = -0.06582681	
)

stats$estimate
```



The final planned contrast compares the sustainer ratings for the second cause in "irreversible" and "reversible". It must be looked up from the long list of contrasts below: 

```{r}
contrasts <- emmeans(a1, ~ entity*structure)
s <- pairs(contrasts, adjust = "none") 


s
confint(s, level = 0.95)
```

The relevant contrast is: "second..sustained stable - second..sustained unstable", which is significant. 

Get effect size: 

```{r}
#structure    entity             lsmean    SE df lower.CL upper.CL
# unstable effect second..sustained   9.34 0.454 98   8.4397    10.24
# stable effect   second..sustained   3.34 0.454 98   2.4397     4.24


#SDs
# unstable effect	second \nsustained	2.0564483	
# stable effect	second \nsustained	4.0436900


stats <- d.ind.t(
  m1 = 9.34,
  m2 = 3.34,
  sd1 = 2.0564483	,
  sd2 = 4.0436900,
  n1 = 50,
  n2 = 50,
  a = 0.05
)

stats$estimate
```

